---
title: "Progression in mFARS/SARA/ICARS"
author: "Christian Rummey"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
    number_sections: yes
  html_document:
    toc: yes
    df_print: paged
    number_sections: yes
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 100
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=9, fig.height=4, fig.align='center', echo=F, warning=F, message=F, cache=F)

# parlist <- c('mFARS','SARA','ICARS','FARS.E','SARA.ax','ICARS.ax','FARS.BC','FARS.B','SARA.ki','SARA.ku' ,'ICARS.ki')
# age.grps <- c(0,8,16,25,40,100)
# age.lbls <- c('<8y', '8-15y', '16-24y', '25-40y', '>40y' )

parlist <- c('mFARS','SARA','ICARS','FARS.E','SARA.ax','ICARS.ax','FARS.BC','SARA.ki','ICARS.ki')

age.grps  <-  c(0,8,12,16,25,40,100)
age.lbls <- c('<8y', '8-11y', '12-15y', '16-24y', '25-40y', '>40y' )

dt. <- readRDS('DATA derived/long.data.rds') %>% 
  mutate(paramcd = factor(paramcd, parlist)) %>% 
  # filter(forslope == 1) %>% 
  # select(-c('forslope', 'int', 'chg', 'dev.y')) %>% 
  droplevels()

dt. %<>% 
  left_join(.rt('../DATA other/scales.txt') %>% select(paramcd, maxscore)) %>% 
  mutate(paramcd = factor(paramcd, parlist)) %>%
  mutate( age.grp = cut(age, age.grps, labels = age.lbls, right = T))

```

# Methods

- General methodology follows the longitudinal changes analyses (latest paper)
- Scores were analyzed in parallel, both on the absolute scale, and using the percentage of total  
- \% of Total Changes were added for Comparison
- y-axis of percentual Changes and SRMs were capped (20% and 1.2, respectively, for better comparison)

# Results

## Total mFARS, SARA, ICARS

```{r}

scores.list <- c('mFARS','SARA','ICARS')

tmp.A <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  mutate ( chg = 100*chg/maxscore, aval = 100*aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

tmp.B <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  # mutate ( chg = chg/maxscore, aval = aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

label.y.A <- -max(tmp.A$m)*.1
label.y.B <- -max(tmp.B$m)*.1

A <- tmp.B %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.B), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_tr+
  ggtitle('Changes in Total Scores (absolute)')+
  ylab('Mean Change')+
  xlab('Age Group')

B <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.A), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('Changes in Total Scores (percentage)')+
  ylab('Mean Percentage Change')+
  xlab('Age Group')


C <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = SRM)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('SRMs')+
  xlab('SRM')+
  xlab('Age Group')

A
B + coord_cartesian(ylim = c(  0,  20))
C + coord_cartesian(ylim = c(-.1, 1.2))

```

## Axial Function, mFARS, SARA, ICARS

```{r}

scores.list <- c('FARS.E','SARA.ax','ICARS.ax')

tmp.A <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  mutate ( chg = 100*chg/maxscore, aval = 100*aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

tmp.B <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  # mutate ( chg = chg/maxscore, aval = aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

label.y.A <- -max(tmp.A$m)*.1
label.y.B <- -max(tmp.B$m)*.1

A <- tmp.B %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.B), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_tr+
  ggtitle('Changes in Total Scores (absolute)')+
  ylab('Mean Change')+
  xlab('Age Group')

B <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.A), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('Changes in Total Scores (percentage)')+
  ylab('Mean Percentage Change')+
  xlab('Age Group')


C <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = SRM)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('SRMs')+
  xlab('SRM')+
  xlab('Age Group')


A
B + coord_cartesian(ylim = c(  0,  20))
C + coord_cartesian(ylim = c(-.1, 1.2))

```


## Appendicular Function, mFARS, SARA, ICARS

```{r}

scores.list <- c('FARS.BC','SARA.ki','ICARS.ki')

tmp.A <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  mutate ( chg = 100*chg/maxscore, aval = 100*aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

tmp.B <- dt. %>% 
  filter ( !is.na(int) ) %>% 
  filter ( int == '1y' ) %>% 
  # mutate ( chg = chg/maxscore, aval = aval/maxscore ) %>%
  group_by(paramcd, age.grp, amb, int ) %>% 
  filter ( paramcd %in% scores.list ) %>%
  summarise(n = n(), m = mean(chg), sd = sd(chg), SRM = m/sd) 

label.y.A <- -max(tmp.A$m)*.1
label.y.B <- -max(tmp.B$m)*.1

A <- tmp.B %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.B), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_tr+
  ggtitle('Changes in Total Scores (absolute)')+
  ylab('Mean Change')+
  xlab('Age Group')

B <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = m)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_text(aes(label=n, y = label.y.A), position = position_dodge(width = .6), size = 2)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('Changes in Total Scores (percentage)')+
  ylab('Mean Percentage Change')+
  xlab('Age Group')


C <- tmp.A %>% 
  ggplot()+geom_col(position = position_dodge(width = .6), alpha = .8)+
  aes(x = age.grp, y = SRM)+
  aes(fill = paramcd)+
  facet_grid(.~amb)+
  geom_hline(yintercept = 0, linetype = 'dotted')+
  .leg_none+
  ggtitle('SRMs')+
  xlab('SRM')+
  xlab('Age Group')


A
B + coord_cartesian(ylim = c(  0,  20))
C + coord_cartesian(ylim = c(-.1, 1.2))

```

